---
title: "Project2"
format: html
editor: visual
---


## Preparation for App
reading in Seoul Bike sharing data for the year-long period from December 2017 through November 2018. Locale argument was included to allow data to read in properly, as the formatting for date seemed to be giving it trouble.
```{r}
library(tidyverse)
bike <- read_csv("SeoulBikeData.csv", locale=locale(encoding="latin1"))
```

One-way and Two-way contingency tables. Due to the nature of bike count being a numeric variable with many possible values, I instead created a range of values for ease of viewing the table.
```{r}
holiday <- bike |> group_by(Holiday) |> summarize(count = n())
holiday

hourly <- bike |>
  mutate(
    count_range = case_when(
      `Rented Bike Count` < 200 ~ "0–199",
      `Rented Bike Count` < 400 ~ "200–399",
      `Rented Bike Count` < 600 ~ "400–599",
      `Rented Bike Count` < 800 ~ "600-799",
      `Rented Bike Count` < 1000 ~ "800-999",
      `Rented Bike Count` < 1200 ~ "1000-1199", 
      TRUE ~ "1200+"
    )
  ) |>
  group_by(Hour, count_range) |>
  summarize(count = n(), .groups = "drop") |>
  arrange(Hour, count_range)

hourly
```

Numerical summaries for 
```{r}

```
